---
title: "Market neutral strategy - SnP 500 and Berkshire Hathaway"
author: "Jirong"
date: 2018-08-24
categories: ["investment", "finance"]
tags: ["finance", "investment"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Market neutral strategy

As the negative news pile up (trade wars, slump in economy growths, etc), I sought for market neutral stategies that could perform well in any market environment.

An idea that struck me recently is to exploit the pair between Berkshire and SnP 500.

The SnP500/ Berkshire ratio has been falling over the years - insinuating that Berkshire has lost its magic of outperforming the index over a couple of decades.

That being said, it's still widely regarded as a safe haven in times of trouble.

### Strategy

As the market is facing headwind and I sought for a market neutral strategy. Here's what I did,

- I derive the SnP500/ Berkshire ratio
- And I fix a Bollinger Band around the ratio with n = 200 days as the moving average parameter - with 2 SD as the lower (lb) and upper bound (ub) line. Bollinger band ratio accounts for the downward trend in the ratio over time.
- If the ratio is below the lb, I will long SnP500 and short Berkshire Hathaway. And when it mean reverts and touches the middle moving average, I will exit the positions
- Conversely if the ratio is above the ub, I will short SnP500 and long Berkshire Hathaway. Similarly, I will exit the position when it touches the moving average 

### Results

So how did the strategy fare? Fairly impressive I must say.

Sharpe ratio is around 0.8. Annualized return is around 5.7% with the positions only in the market 30% of the time!

That being said, I'm cherry picking here because the performance before this period is sub-par; probably because of a change in market regime. You may execute my code to stress-test this simple strategy.

### Disclosure 
I may execute this strategy in the near term.

## Running packages

```{r pressure, echo=FALSE}
#Note: Research on Bollinger bands - Pair Trading Ratio
suppressMessages(sapply(c("zoo", "tidyr", "plyr", "dplyr",
         "gtools","googlesheets", "quantmod", 
         "urca", "PerformanceAnalytics", "parallel", "TTR"), require, character.only = T))

source('util/calculateReturns.R')
source('util/calculateMaxDD.R')
source('util/backshift.R')
source('util/extract_stock_prices.R')
source('util/cointegration_pair.R')

```

## Function
```{r}
#Function using bollinger band
lf_bollinger_pair_trading = function(stock1, stock2, start_date, end_date, prop_res, bband_days){
  
#Start of function
data1 = df_crawl_time_series(stock1, start_date, end_date)
data1 = subset(data1, select = c("Date", "Open", "Adj.Close"))
names(data1) = c("Date", "Open", "Close")
data1$Date = as.Date(data1$Date)

data2 = df_crawl_time_series(stock2, start_date, end_date)
data2 = subset(data2, select = c("Date", "Open", "Adj.Close"))
names(data2) = c("Date", "Open", "Close")
data2$Date = as.Date(data2$Date)

#Training and testing index
data1 = xts(data1[, -1], order.by = data1[, 1])
data2 = xts(data2[, -1], order.by = data2[, 1])

data = merge(data1, data2)
data = as.data.frame(data)
data = subset(data, !is.na(data$Close) & !is.na(data$Close.1))

data$ratio = data$Close/ data$Close.1

# plot(data$ratio)
bb_ratio = data.frame(BBands( data$ratio, n = bband_days))
data = cbind(data, bb_ratio)
data_sub = tail(data, round(nrow(data) * prop_res, 0))

plot(data_sub$ratio)
lines(data_sub$mavg, col = "red")
lines(data_sub$up, col = "blue")
lines(data_sub$dn, col = "green")

#If lower than 
data_sub$longs <- data_sub$ratio <= data_sub$dn # buy spread when its value drops below 2 standard deviations.
data_sub$shorts <- data_sub$ratio >= data_sub$up # short spread when its value rises above 2 standard deviations.

#  exit any spread position when its value is at moving average
data_sub$longExits   <- data_sub$ratio >= data_sub$mavg
data_sub$shortExits <- data_sub$ratio <= data_sub$mavg


# #  define indices for training and test sets
# trainset <- 1:as.integer(nrow(data) * prop_train)
# testset <- (length(trainset)+1):nrow(data)

#Signal
data_sub$posL1 = NA
data_sub$posL2 = NA
data_sub$posS1 = NA
data_sub$posS2 = NA

# initialize to 0
data_sub$posL1[1] <- 0; data_sub$posL2[1] <- 0
data_sub$posS1[1] <- 0; data_sub$posS2[1] <- 0

data_sub$posL1[data_sub$longs] <- 1
data_sub$posL2[data_sub$longs] <- -1

data_sub$posS1[data_sub$shorts] <- -1
data_sub$posS2[data_sub$shorts] <- 1

data_sub$posL1[data_sub$longExits] <- 0
data_sub$posL2[data_sub$longExits] <- 0
data_sub$posS1[data_sub$shortExits] <- 0
data_sub$posS2[data_sub$shortExits] <- 0

#positions
data_sub$posL1 <- zoo::na.locf(data_sub$posL1); data_sub$posL2 <- zoo::na.locf(data_sub$posL2)
data_sub$posS1 <- zoo::na.locf(data_sub$posS1); data_sub$posS2 <- zoo::na.locf(data_sub$posS2)
data_sub$position1 <- data_sub$posL1 + data_sub$posS1
data_sub$position2 <- data_sub$posL2 + data_sub$posS2

#Returns
data_sub$dailyret1 <- ROC(data_sub$Close) #  last row is [385,] -0.0122636689 -0.0140365802
data_sub$dailyret2 <- ROC(data_sub$Close.1) #  last row is [385,] -0.0122636689 -0.0140365802

#Backshifting here. But signal is for following day returns!. So can still use latest Z-score
data_sub$date = as.Date(row.names(data_sub))
data_sub = xts(data_sub[,-which(names(data_sub) == "date")], order.by = data_sub[, which(names(data_sub) == "date")])

#Doesn't account for number of shares!!!!!
data_sub$pnl = lag(data_sub$position1, 1) * data_sub$dailyret1  + lag(data_sub$position2, 1) * data_sub$dailyret2

#Performance analytics
tryCatch({
  # charts_perf = charts.PerformanceSummary(data_sub$pnl)
  charts.PerformanceSummary(data_sub$pnl)
}, error = function(e){})

dd = table.Drawdowns(data_sub$pnl)
ds_risk = table.DownsideRisk(data_sub$pnl)
ret = table.AnnualizedReturns(data_sub$pnl)


df_ret = list(data_sub = data_sub,
              dd = dd,
              ds_risk = ds_risk,
              ret = ret
              )

return(df_ret)

}
```

```{r}
#Parameters to be includded in function
stock1 = "^GSPC"
stock2 = "BRK-B"

start_date = "2000-07-01"
end_date = "2019-12-30"

prop_res = 0.25     #Proportion of results to show
bband_days = 200   #Bollinger band of ratio

#Storing result to function
res = lf_bollinger_pair_trading(stock1, stock2, start_date, end_date, prop_res, bband_days)
```

## Displaying of results

- Drawdown period
```{r}
res$dd
```

- Downside risk
```{r}
res$ds_risk
```

- Returns
```{r}
res$ret
```



